---
title: "Great Salt Lake Rayshader"
author: "Josh Gladwell"
date: "2023-04-20"
output: html_document
---

# Introduction
In this notebook, I will be drafting my code to create a 3d model of the Great Salt Lake in rayshader. I am following [this tutorial](https://www.tylermw.com/a-step-by-step-guide-to-making-3d-maps-with-satellite-imagery-in-r/) to create the model.

Start by importing libraries:
```{r setup, message=FALSE}
library(rayshader)
library(sp)
library(raster)
library(scales)
library(rgdal)
library(magick)
```

Read in the data:
```{r}
gsl_elevation <- raster::merge(
  raster("../../data/raw/SRTM/N40W112.hgt"),
  raster("../../data/raw/SRTM/N40W113.hgt"),
  raster("../../data/raw/SRTM/N40W114.hgt"),
  raster("../../data/raw/SRTM/N41W112.hgt"),
  raster("../../data/raw/SRTM/N41W113.hgt"),
  raster("../../data/raw/SRTM/N41W114.hgt")
)

height_shade(raster_to_matrix(gsl_elevation)) %>%
  plot_map()
```

Now process the satellite imagery:

```{r}
# Rasterize and stack "image A"
gsl_a_rbg = stack(
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038031_20220602_20220609_02_T1/LC08_L1TP_038031_20220602_20220609_02_T1_B4.TIF"), # R
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038031_20220602_20220609_02_T1/LC08_L1TP_038031_20220602_20220609_02_T1_B3.TIF"), # B
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038031_20220602_20220609_02_T1/LC08_L1TP_038031_20220602_20220609_02_T1_B2.TIF") # G
)

# Rasterize and stack "image B"
gsl_b_rbg = stack(
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038032_20220602_20220609_02_T1/LC08_L1TP_038032_20220602_20220609_02_T1_B4.TIF"), # R
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038032_20220602_20220609_02_T1/LC08_L1TP_038032_20220602_20220609_02_T1_B3.TIF"), # B
  raster("../../data/raw/USGS_Landsat/LC08_L1TP_038032_20220602_20220609_02_T1/LC08_L1TP_038032_20220602_20220609_02_T1_B2.TIF") # G
)

# Rasterize and stack "image C"
gsl_c_rbg = stack(
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039031_20220601_20230407_02_T1/LC09_L1TP_039031_20220601_20230407_02_T1_B4.TIF"), # R
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039031_20220601_20230407_02_T1/LC09_L1TP_039031_20220601_20230407_02_T1_B3.TIF"), # B
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039031_20220601_20230407_02_T1/LC09_L1TP_039031_20220601_20230407_02_T1_B4.TIF") # G
)

# Rasterize and stack "image D"
gsl_d_rbg = stack(
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039032_20220601_20230407_02_T1/LC09_L1TP_039032_20220601_20230407_02_T1_B4.TIF"), # R
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039032_20220601_20230407_02_T1/LC09_L1TP_039032_20220601_20230407_02_T1_B3.TIF"), # B
  raster("../../data/raw/USGS_Landsat/LC09_L1TP_039032_20220601_20230407_02_T1/LC09_L1TP_039032_20220601_20230407_02_T1_B2.TIF") # G
)

# Merge images into one stack
gsl_rbg = merge(gsl_a_rbg, gsl_b_rbg, gsl_c_rbg, gsl_d_rbg)

# Plot merged rasterstack
plotRGB(gsl_rbg, scale=255^2)
```

Apply gamma correction to make imagery brighter:
```{r}
gsl_rbg_corrected = sqrt(gsl_rbg)

plotRGB(gsl_rbg_corrected)
```

Now we attempt to combine the satellite imagery with the terrain data:
```{r}
gsl_elevation_utm = raster::projectRaster(gsl_elevation, crs = crs(gsl_rbg), method = "bilinear")
crs(gsl_elevation_utm)

# Save these rasters
writeRaster(
  gsl_elevation_utm, 
  "../../data/modified/rasters/gsl_elevation_utm",
  overwrite=TRUE
)

writeRaster(
  gsl_rbg_corrected,
  "../../data/modified/rasters/gsl_rbg",
  overwrite=TRUE
)

# Try writing as geoTIFF
writeRaster(
  gsl_rbg_corrected,
  "../../data/modified/rasters/gsl_rbg",
  form = "GTiff",
  overwrite=TRUE
)

# # To read it back in:
# gsl_rbg_new <- brick("../../data/modified/rasters/gsl_rbg.gri")
# 
# # Try reading back in geotif (see if it's the same)
# gsl_rbg_tif_new <- brick("../../data/modified/rasters/gsl_rbg.tif")
# 
# plotRGB(gsl_rbg_tif_new)
```

Next step is to crop the map used by defining the desired extent:
```{r}
bottom_left = c(y=-113.307704, x=40.527373)
top_right = c(y=-111.472865, x=41.830442)

extent_latlong = sp::SpatialPoints(rbind(bottom_left, top_right), proj4string=sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
extent_utm = sp::spTransform(extent_latlong, raster::crs(gsl_elevation_utm))

e = raster::extent(extent_utm)
e
```

Now we actually crop the elevation and satellite imagery:
```{r}
gsl_rgb_cropped = raster::crop(gsl_rbg_corrected, e)
elevation_cropped = raster::crop(gsl_elevation_utm, e)

names(gsl_rgb_cropped) = c("r","g","b")

gsl_r_cropped = rayshader::raster_to_matrix(gsl_rgb_cropped$r)
gsl_g_cropped = rayshader::raster_to_matrix(gsl_rgb_cropped$g)
gsl_b_cropped = rayshader::raster_to_matrix(gsl_rgb_cropped$b)

gsl_el_matrix = rayshader::raster_to_matrix(elevation_cropped)

gsl_rgb_array = array(0,dim=c(nrow(gsl_r_cropped),ncol(gsl_r_cropped),3))

gsl_rgb_array[,,1] = gsl_r_cropped/255 # Red layer
gsl_rgb_array[,,2] = gsl_g_cropped/255 # Green layer
gsl_rgb_array[,,3] = gsl_b_cropped/255 # Blue layer

gsl_rgb_array = aperm(gsl_rgb_array, c(2,1,3))

plot_map(gsl_rgb_array)
```

We now introduce scaling into the model:
```{r}
gsl_rgb_contrast = scales::rescale(gsl_rgb_array,to=c(0,1))

plot_map(gsl_rgb_contrast)
```

Finally, we plot the 3d model:
```{r}
plot_3d(
  gsl_rgb_contrast, gsl_el_matrix, 
  windowsize = c(1920,1080), 
  zscale = 15, 
  shadowdepth = -50,
  zoom=0.65, 
  phi=30,
  theta=0,
  fov=70, 
  background = "white", 
  shadowcolor = "#523E2B"
)
render_snapshot(
  title_color = "white", 
  title_bar_alpha = 1
)
```

Turn it into a video:
```{r, warning=FALSE}
angles= seq(0,360,length.out = 1441)[-1]
for(i in 1:1440) {
  render_camera(theta=0+angles[i])
  render_snapshot(filename = sprintf("imgs/gsl%i.png", i))
}
rgl::rgl.close()


system("ffmpeg -framerate 60 -i ./imgs/gsl%d.png -pix_fmt yuv420p vid/gsl.mp4")
```



